Supporting Information for\\ ``A weakly structured stem for human origins in Africa''

### 2022-12-27 supplement.tex
refs for fixing dates

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
and the inferred split between East and West African populations occurred about
$10$ka earlier. These recent split dates are at odds with archaeological
evidence for an earlier Out-of-Africa (OOA) expansion
\citep[e.g.]{Hublin2020-nk,Hajdinjak2021-jo}, with the estimated date for
Neanderthal admixture into Eurasian populations \citep[e.g.][]{Reilly2022-ym},
and with previous genetics-based dating of splits between African and Eurasian
populations \citep[e.g.]{Gutenkunst2009-ff,Jouganous2017-pq,Kamm2020-vn}.  The
parameter uncertainty of our inferred divergence times (which accounts for the
non-independence of sites) does not account for these differences: in the
high-likelihood models, the $95\%$ confidence intervals for the East
African-European split was roughly $38 \pm 5$ka, and for the East/West African
split was $45 \pm 10$ka, with the uncertainties reflecting 2 standard
deviations according to bootstrap of genomic regions. We therefore expect these
discrepancies to result from model misspecification in the details of the
divergence between the ancestors of OOA and East African populations.

Our models suppose a single split leading to the OOA event followed by constant
symmetric migrations between eastern African and OOA populations, which is an
oversimplification. Previous studies that have focused on reconstructing the
OOA divergence time and subsequent migration rates typically included more
Eurasian populations and multiple epochs that allow for differing continuous
migration rates and population sizes. Since a detailed model for the OOA (and
Back-to-Africa) event was not our main focus, we chose to keep this simplified
model, but fix the east African-Eurasian split at $50$ka \citep{Hublin2020-nk}.
This also allowed us to include Neanderthal admixture into the OOA population
at $45$ka.

>>>>

In our preliminary models, the unconstrained inferred split times between East
African populations and Europeans was inferred to have occurred around $40$ka,
and the inferred split between East and West African populations occurred about
$10$ka earlier. These recent split dates are at odds with archaeological
evidence for an earlier Out-of-Africa (OOA) expansion
\citep[e.g.]{Fu2014-am,Hublin2020-nk,Hajdinjak2021-jo}, with the estimated date
for Neanderthal admixture into Eurasian populations
\citep[e.g.][]{Reilly2022-ym}, and with previous genetics-based dating of
splits between African and Eurasian populations
\citep[e.g.]{Gutenkunst2009-ff,Jouganous2017-pq,Kamm2020-vn}. The parameter
uncertainty of our inferred divergence times (which accounts for the
non-independence of sites) does not account for these differences: in the
high-likelihood models, the $95\%$ confidence intervals for the East
African-European split was roughly $38 \pm 5$ka, and for the East/West African
split was $45 \pm 10$ka, with the uncertainties reflecting 2 standard
deviations according to bootstrap of genomic regions. We therefore expect these
discrepancies to result from model misspecification in the details of the
divergence between the ancestors of OOA and East African populations.

Our models suppose a single split leading to the OOA event followed by constant
symmetric migrations between eastern African and OOA populations, which is an
oversimplification. Previous studies that have focused on reconstructing the
OOA divergence time and subsequent migration rates typically included more
Eurasian populations and multiple epochs that allow for differing continuous
migration rates and population sizes. Since a detailed model for the OOA (and
Back-to-Africa) event was not our main focus, we chose to keep this simplified
model, but fix the East African-Eurasian split at $50$ka
\citep{Fewlass2020-se,Hublin2020-nk}. This also allowed us to include
Neanderthal admixture into the OOA population at $45$ka.
